---
title: "LeMANS (Length-based Multispecies Analysis by Numerical Simulation)"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{LeMANS}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE,   comment = "#>" ,fig.width = 8, fig.height = 8,fig.align="center")
library(LeMANS)
```


The LeMANS package is an implementation of the LeMANS model as described in the Hall et al (2006) and Rochet et al (2011) [publications](#pubs). The data provided (bundled with the package - lazily loaded) reflects the Georges Bank (USA) data described in the Rochet et al. paper.

* `rochet_GB_modelSetup`
* `rochet_GB_parameterValues`
* `rochet_GB_initialValues`
* `rochet_GB_foodweb`
* `rochet_GB_species`


The contents of these can be viewed simply by typing in the names in the console. To view the description of a data object use help. For example `?rochet_GB_parameterValues`.

## Usage

To run the model right out of the box:
```{r run1, echo=TRUE}
    results <- key_run(Ffull=0.4, nYrs=50, rochet_GB_modelSetup, rochet_GB_parameterValues, rochet_GB_initialValues, rochet_GB_foodweb, rochet_GB_species)

```


`results` is a list of several output variables (catch, recruits, M1, M2, F, spawning stock biomass, N, and more).

```{r run2, echo=TRUE}
    str(results)
```

[Plotting](#plotting) routines help visualize these output variables

Type `?key_run` for more help 

## Intermediate functions

Several functions are invoked when running the model. These functions can be changed by the user to suit their needs.

* `calc_M1`
* `calc_M2_c`
* `calc_maturity`
* `calc_phi`
* `calc_population_growth`
* `calc_ration`
* `calc_recruits`
* `calc_sizePref_suitability`
* `calc_F`

Help is available for all functions or feel free to write your own to replace any of these.

## Plotting {#plotting}

Visualization of output is critical. A basic plotting tool is provided to view output. The user is encouraged to write their own plotting routines.

###To produce a plot of catch by species for each size class with all of the species yaxes scaled to their output:
```{r plot_c, echo = TRUE}
    plot_key_run(results$catch/1E6, ylabel = "Catch (millions individuals)", speciesNames=rochet_GB_species, scales="free")

```


###To produce a plot of total abundance aggregated over each size class with species yaxes fixed across all species:
```{r plot_n, echo = TRUE}
    plot_key_run(results$N/1E6, ylabel = "Abundance (millions individuals)", is.aggregated=T, speciesNames=rochet_GB_species, scales="free")

```


###To produce a plot of M2 predation for each size class with species yaxes fixed across all species:
```{r plot_m2, echo = TRUE}
    plot_key_run(results$M2, ylabel = "Predation mortality", speciesNames=rochet_GB_species, scales="fixed")

```

###To produce a plot of the suitability forage fish to other species in the model:
```{r plot_suit, echo = TRUE}
    plot_key_run(results$suitability, ylabel = "Suitability", speciesNames = rochet_GB_species, scales="free",speciesSuitability = "Forage fish",predOrPrey = "prey")

```


## Publications {#pubs}

* Hall et al. (2006). A length-based multispecies model for evaluating community responses to fishing. Can. J. Fish. Aquat. Sci. 63:1344-1359.

 * Rochet et al. (2011). Does selective fishing conserve community biodiversity? Prediction from a length-based multispecies model. Can. J. Fish. Aquat. Sci. 68:469-486


